Autonomous Oceanographic Data Collection. Learning Excursion Sets of Vector-valued Gaussian Random Fields \\or\\ Learning Excursion Sets of Vector-valued Gaussian Random Fields for Autonomous Ocean Sampling

### 2020-06-26 ans_to_reviewers/answers-to-reviewers.tex

%\begin{itemize}[noitemsep,topsep=0pt,parsep=0pt,partopsep=0pt]

%\item[1.1.1]\label{ec1} 
\vspace{5mm}
\noindent \aecom \textbf{1}:

General: The inferential problem considered in the paper is that of estimating the “boundary” of a plume. However, the authors never define exactly what a “boundary” is. Such a definition should be provided, as well as some discussion of why determining the location of a boundary is so important. Why is it not sufficient to create a map of predictions (with an accompanying map of prediction error variances) for the variable(s) under study throughout the region of interest (a problem for which there is a huge literature on sampling design)?\par

\reply: In the revised version we have strengthened the motivation for focusing on ESs and using the EIBV criteria. The revised version has an entirely rewritten Introduction, and Section 2 now includes general background for random sets and excursion sets. (We added references, for instance \cite{Molchanov2005} on random sets.) The revised version provides results that go beyond the EIBV (with proofs in the new Appendix), but the EIBV is used in experiments because it is computationally tractable onboard the AUV.

@@

\vspace{1em}

%\item[1.1.2]\label{ec2} 
\vspace{5mm}
\noindent \aecom \textbf{2}:

Page 8, second paragraph of Section 4.1: The assumptions of separability and isotropy, and even stationarity, are not really necessary for the methodological development of Sections 4–6. I suggest that you wait to make these assumptions until you get to the simulation studies and example.\par

\reply: We agree that the detailed statistical modeling assumptions are not necessary for the methodological contributions, so we now bring up this model specification when needed for the examples. 
In the open-source code that is now available, we also implement non-stationary and anisotropic covariance models. These represent situations that have some relevance for our river plume case. The code we post along with the paper facilitate testing so users can input more complex models to see how it influences designs.

@@

\vspace{1em}

%\item[1.1.6]\label{ec1} 
\vspace{5mm}
\noindent \aecom \textbf{6}:

Page 18, first paragraph of Section 7.1: The numerical values of the regression parameters given here are not meaningful without knowing the units of “West.”\par

\reply: We clarified the parameter specification.
In practice the trend captures the temperature and salinity levels on the domain defined by the waypoint graph which is constructed to cover both riverine water and ocean water. For the field experiment (Section 4) we clarified this specification.

@@

%\item[1.2.3]\label{emc3} 
\item Page 19, fourth line from bottom: Change ``as also" to ``as well as"

\reply: We have changed this in the text.

%\item[1.2.4]\label{emc4} 
\item Page 30, second and third references listed: ``et al." is listed after the authors’ names for each
of these two references, but actually there are no additional authors beyond the two listed.

\reply: We have made this change to the citations: \cite{genton2015cross} and \cite{french2013spatio}.
\end{itemize}

@@

\reply: Thank you.

\item{\revcom  :}\label{r1c3}

The things that stand out as novel in this paper are: the need for a sequential design for driving an autonomous under water vehicle; a way of characterizing water plumes that are washed down from the land into the sea. Papers co-authored by one of the authors (Ginsbourger) of this paper inevitably means some overlap of ideas with the other papers, although the two step ahead greedy design algorithm seems novel. I seem to recall that Jon Lee published that algorithm albeit for determining the maximum entropy optimum by finding optimal points sequentially over space not time as in this paper. The literature on spatial design is huge and often involves spatial sequential designs. The handbook chapters by Zidek and Zimmerman (or is the other around) might be worth citing for the benefit of readers inter- ested in some of the background.

\reply: We have improved and augmented the discussion on other design criteria and why the ES related EIBV criterion is used in our case. The paper also contains statistical results that go beyond this criterion (with proofs in the new Appendix), but the EIBV is used in practice because it is computationally tractable onboard the AUV. In this part, we cite the handbook reference on design \cite{zidek2019monitoring} and the paper we believe that contains a comment from Jon Lee \citep{bueso1998state}.

@@

\reply: We have restructured the paper. See Point 1. at the start of this reply letter.

\item{\revcom :}\label{r1c6}

It is not clear why the binary classification and excursion sets are being used. Is there really a sharp boundary at the edge of a plume? Why not mapping the salinity and temperature fields instead? After all, on page 19, “achieving good predictive performance” is mentioned as a desirable feature of a good design. A lot of work has been done on that topic. And what about indicator kriging (along with logit regression) that generalizes indicator kriging that has been used to identify plumes? Modelling the p(x) is more like prohibit analysis-what is the advantage.

\reply: We added reference \cite{Molchanov2005} on random sets and motivated this topic for our application and more general settings. See our general Point 2. and 3. at the start of this reply letter.

@@

\reply: The design criteria for multivariate random fields are new. We further improved our statistical contribution by developing more general results (with proofs in Appendix). There are finally new elements in tailoring this sequential design methods in the context of the AUV application with its computational and navigation limitations. See our general Point 2 and 3 on top.

\item{\revcom  :}\label{r1c8}

While I can see some advantages of simplicity in the IBV of Bect et al for binary data, one could have used entropy as it has information-theoretical credentials. Furthermore it could cover design both for spatial mapping as well as indicator mapping. Finally model uncertainty could be included in the overall assessment of uncertainty. The more general issue here is what objective function should be optimized in the design? Are we in a multi-attribute criterion situation?

\reply: The traditional entropy calculations for this Gaussian Random Field model aim to reduce the log determinant of the updated covariance matrix at the grid locations. It will not vary with the data realizations, so in our setting this information criteria will always give the same path. The EIBV criterion is aligned with our ES motivation for this study. When focusing on the expected EIBV, the paths will vary with the data as we see in our simulation study and in the real-data example. We see this as a clear benefit of the approach compared with entropy-based calculations. Still, sampling approaches using hybrid criteria or multi-attribute criteria might balance goals in such situations, but we leave that for future research. (This is mentioned as further work in Closing remarks.)

@@

\item{\revcom  :}\label{r1c9}

On the issue of covariance separability, objections have always been raised when I have used it as its unrealistic. My rejoinder has always worked, that of citing the “all models are wrong”, George Box’s famous phrase. But what I was able to show in each case is a cross-validatory assessment that showed the model was “useful”. I wonder if such a model assessment could be made here to support the assumption of separability and other things as well. By the way, in spatial mapping at least, point predictors are pretty robust against mis-specification of the covariance. Where things go sideways is in the confidence bands or ellipsoids they produce. What are the error bands when modelling binary responses? Surely the predicted plumes have fuzzy boundaries.

\reply: At this spatial scale there is often not so much data to infer a detailed statistical model, and this is the reason for our parsimonious model in the real-data case. In other related situations, for instance that of larger fjord or ocean fronts, which are at a larger scale, there would usually be more data available for instance in the form of satellite data or from computer model simulations. 
That said, we do provide some validation of the model via residual analysis. Moreover, the main contribution of this paper is on design more than modeling. Finally, the online resources with open-source code have demos using more complex covariance models (non-stationary variance and anisotropic random fields).

@@

We agree that more complex models would matter more if we step up to a criterion such as volume calculations.

\item{\revcom  :}\label{r1c10}

The strategy at the bottom of page 17 seems ad hoc. Could you not use the data collected with the two ahead design to get better estimates of parameters and then with these parameters and thereafter a better myopic design?

\reply: We have chosen not to explore any re-estimation of parameters. The resources onboard the AUV do not allow such computationally demanding evaluations. Overall, we believe the capabilities and limitations of the AUV are clarified in the revised version of the paper.

@@

\item{\revcom :}\label{r1c11}

Are these results generalizable? In other words, how much plume-to-plume variability is there? One of the problems with modelling using a lot of parameters, is there instability over replicate experiments. To put it another way which model parameters represent plume-population parameters and which represent random plume effects?

\reply: We agree that this is a simple model for plume temperature and salinity variation. However, it is not common to have ocean model data or satellite data at this spatial resolution, which would be required for detailed model specification. So rather than overfitting, we have chosen a parsimonious model.

@@

\item{\revcom :}\label{r1c12}

The separability assumption is to be relaxed in future work. Should publication of this case study await the proposed improvement? It would make for a much better paper.

\reply: High-resolution ocean models would be required for such detailed model specification, and it might be doable with large efforts in careful numerical implementation. But we believe that spatio-temporal variations would be a more realistic extension. The methodological approach for sampling designs, which is the main contribution of the current paper, would however still be the transferable.

@@

\item{\revcom : (specific comments)}\label{q11}

\begin{itemize}[noitemsep,topsep=0pt,parsep=0pt,partopsep=0pt]

\item “joint excursion sets” in the title would be a turnoff for most statisticians. Having read the paper pretty carefully I still am not sure why they are so labelled.\par
\reply: We avoid the term' joint excursion sets' in the revised version. The revised version build on a more general notation around random sets and EIBV for multivariate random fields, but we link to the ES for temperature and salinity whenever it is appropriate to exemplify concepts and show results.

@@

\vspace{1em}

\item Why on page 13 were those particular values chosen for the means $\mu$? Would a different choice have made a difference?\par

\reply: A preliminary run was used to set these trends. We have clarified the description in the paper. At a larger scale, one could here use ocean model inputs or satellite data.

@@

\reply: We removed this Figure in the revision.

\vspace{1em}

\item Why are the ES’s called ”excursion sets? Why excursion probabilities?\par

\reply: The revised Section 2 contains a general formulation for random sets, which is then exemplified by our ESs for the multivariate process. In the Introduction and Section 2 we motivate the ES for multivariate processes for this application. In many applications it is important to find zones with joint high values of many variables. In our context, the temperature and salinity are proxies for biological variables which are not easily measured.

@@

\item I did not spot a definition for the “waypoint graph” a phrase used several times in the paper. Again points to the need to identify the audience. I doubt that most statisticians would know that phrase.\par
\reply: The waypoint graph is clarified in the revised version.  
\vspace{1em}

\item In statistics journals, random variables are commonly labelled by Roman characters e.g. Z, parameters in Greek e.g. $\xi$. I realize that in other scientific domains that distinction is not always made but I do recommend the change if the paper is to appear in a stats journal.\par

\reply: We use $Z$ for the random field in the revised version.
\vspace{1em}

\item There is not much of novelty in the Gaussian Process theory used in the paper- its pretty much standard. I suggest that the derivations stated in detail be summarized by just the results. Statisticians will know these things and users will not care.\par

\reply: We collapsed the previous Section 3, 4 and 5 to new Section 2 containing the statistical novelty on EIBV closed-form expressions. The proofs are in the new Appendix.

@@

I know very little about the AUVs you discuss, and I suspect many of your readers will be in a similar position. Additional contextual information would be helpful

\begin{itemize}[noitemsep,topsep=0pt,parsep=0pt,partopsep=0pt]

\item[2.1.1] What sort of onboard computing do they have? It may not be something like an Intel 2.1 GHz, 4-thread processor, but a more informative description of something most readers relate to would be helpful. Similar information related to RAM and hard drive space would be helpful. What are the computational limits of these machines?\par

\reply: We have clarified the description about AUVs and their computational capabilities in Section 4. 
It is a multicore GPU NVIDIA Jetson TX1 (quad-core 1.91 GHz 64-bit ARM machine, a 2-MB L2 shared cache, and 4 GB of 1600 MHz DRAM). 
\vspace{1em}

@@

\item[2.1.2] How long do the AUVs typically explore in one go? Is there a minimum/maximum time?\par

\reply: Maximum speed and payload are limiting coverage. For AUV operations in general, this ranges from minutes to 24h for larger crafts. Time constraints for operations are case specific, and also typically related to costs of ship time and man-hours during operation.  
\vspace{1em}

@@

\item[2.1.3] What kind of movements can the AUVs make? Chess-like moves (horizontal, vertical, diagonal)?\par

\reply: The AUV used in our application is free to move in 6 degrees of freedom on the triangulated graph. This is shown in Fig. 4 with the waypoint graph. (They are essentially flying under water. In larger-scale oceanographic applications they also perform yo-yo movements in depth.) 
\vspace{1em}

@@

\item[2.1.4] Is there a minimum/maximum number of points you want to sample?\par

\reply: The coverage available is usually limited by man-hours to follow operations, but for large-time operation also on battery capacity. The important question here is in any case "How do one prioritize the correct points to visit?", which is central to the development of designs, and we believe the EIBV which we develop here is a useful criterion in several settings. 
\vspace{1em}

@@

\item[2.1.5] What is the size of the grid that is typically considered? i.e., what is a typical value of $n$ in Eq.(9)?\par

\reply: The waypoint grid size should be small enough to resolve the process that is of interest and to avoid aliasing. Usually it is in the 100s or 1000s. A guiding rule of thumb provided by the Nyquist theorem is to sample at least twice in time for the shortest significant record period, and twice in space for the shortest significant length to either resolve or eliminate (by filtering) scales of oceanic variability shorter than those being studied. In practice, this means mapping at an adequate spatial resolution faster than significant changes – in the phenomena – occur. 
\vspace{1em}

@@

\item[2.1.6] Does the AUV use a selection algorithm from the very beginning or is there a minimum number of initial measurements that are made before the optimization algorithm begins?\par

\reply: As is common in such field trials with AUVs, we gathered data in an initial run that is also used to test the depth and bearings navigation systems on the AUV. (With the low salinity in the river this is not always trivial as the AUVs can sink.) From this initial data, the model parameters were tuned and the AUV started conducting adaptive sampling using the optimization strategy.
\vspace{1em}

@@

\reply: In the revised version the notation has been improved. There is limited data for training in our case, so there is a linear regression trend while covariance parameters have been set constant across locations.

\item{\revcom :}\label{r2c3}

p. 9, How reasonable is the assumption of stationary and isotropic random fields? You may be limited by the computational abilities of the AUVs so that more complex models are computationally infeasible. That is a valid reason to choose a simpler, less accurate model. I’m just wondering if that is the reason here.

\reply: We agree that this is a simplified model. However, one would need ocean model data or satellite information to fit a more complex model, and this is often not available at our spatial resolution. However, in the accompanying Python code we run cases with more complex covariances (non-stationary variance and anisotropic covariance), so the interested user can test if / how it matters for the design. We believe spatio-temporal models would be a very interesting next step and present that as further work.

@@

\item{\revcom  :}\label{r2c4}

p. 9, Eq. (10). Are the temperature and salinity processes at a specific location effectively independent of one another? (I truly don’t know). That is what is implied by Eq. (10).

\reply: The temperature and salinity are dependent in the Gaussian Random Field field model. But the temperature and salinity data are collected from different instruments, and we assume that they are conditionally independent, given the random field variables. That said, other measurements might be extracted or processed from common sensor readings and would then be conditionally dependent. We mentioned this briefly in the model assumptions descriptions.

@@

\item{\revcom :}\label{r2c5}

p. 9, Eq. (11). The form of the equation is reminiscent of the Sherman-Morrison-Woodbury solution for fixed-rank kriging. $\bSigma$ may not be small here, but I wonder if there may be a computational advantage to considering an alternative computational solution.  See e.g., Cressie, N. and Johannesson, G. (2008), Fixed rank kriging for very large spatial data sets. Journal of the Royal Statistical Society: Series B (Statistical Methodology), 70: 209-226. doi:10.1111/j.1467-9868.2007.00633.x.

\reply: The update formula assimilates relatively modest data sizes at each waypoint. Moreover, only the near-surface grid is considered here (two dimensional spatial random field model). This means that the computational requirements are mainly driven by the number of combinations for sampling designs, not the size of the domain or the data dimension. 
We still agree that computational approximations, such as fixed-rank-kriging, would be useful for larger-size three dimensional applications.

@@

\item{\revcom  :}\label{r2c6}

p. 9, Eq. (13). Why does $\bm$ have a distribution before seeing the data? I can see how $\bm(\by)$ would have a distribution since it depends on $\by$. It almost feels Bayesian, but nothing indicates that this analysis is being pursued from a Bayesian perspective. Why is it valid for $\bm$ to have a distribution prior to seeing data?

\reply: These parts of the paper are largely rewritten. We have also generalized the closed-form solutions. (See point 2 and 3 at the start of reply letter.) The main body contains results of general interest and provide insight about the statistical novelty. The proofs are in Appendix with a succinct notation.

@@

\reply: We have generalized the closed-form solutions. (See reply above.)

\item{\revcom :}\label{r2c12}

This method relies on computing the marginal probabilities in each grid cell. This is computationally convenient and is likely required because of the limited computing power of the AUVs. However, the optimization should technically be performed with respect to the joint distribution across all grid cells. While the context may not allow this in practice, it would be helpful to get an idea of how much error is introduced by this approximation. I recommend creating a small but realistic-as-feasible simulation study comparing the accuracy of the probabilities, a comparison of the computational expense, and examining if/how navigation decisions change based on whether one uses the marginal/joint computations.

\reply: In the revised version we provide closed-form solutions to more general criteria (beyond the EIBV), such as volume calculations involving coupling between locations. But as you mention, the real-world application has computational limitations, and we still use marginal distributions and EIBV in the examples.

@@

\item{\revcom :}\label{r2c13}

Fig. 7 is difficult to understand. The goal is to show the various paths taken by the AUV in the
simulation example. But it is very difficult to determine which paths are very common and which
are uncommon. Some ideas that may make this graphic more useful:

\begin{quote}
-Don’t worry about the paths but focus on the nodes visited. Represent each node using a circle or hexagon.\par
-Make the node shading darker depending on whether that node is visited in a
simulation. Nodes with no visits would be white. Nodes with 100 visits would be black.\par
-The true excursion set is a random variable, but it would be informative to overlay the
average position of the excursion set over each of the panels so the reader can see how
the simulated AUV moves back and forth across the excursion set.
\end{quote}

\reply: We have removed this figure because it confused the explanation. The online resources contain more results, including sampling paths, and this code allows users to generate sampling results by themselves.

@@

\reply: Code is improved in the revision and provided for open-source usage: \url{https://cedrictravelletti.github.io/Submarine/}.

\item{\revcom: Minor comments}\label{r2cminor}

\begin{itemize}[noitemsep,topsep=0pt,parsep=0pt,partopsep=0pt]

\item[2.15.1] “closed-form” is sometimes written as “closed form” (e.g., p. 7, p. 18)

@@

\begin{answers}

\item{\revcom - Major 1:}\label{r3c1}

I think this is an interesting paper. The paper is well written, 
and the authors have motivated the application well. 
The authors indicate several paths forward for work in this area, which I think pose interesting questions. They may want to elaborate a bit more on those.

### 2020-06-26 ans_to_reviewers/answers-to-reviewers.tex

\reply: In the revised 'Closing remarks', we elaborate more on: \\
i) computational approximations required for three-dimensional way-point graphs and space-time modeling including relevant references \citep{sigrist2015stochastic,richardson2017sparsity} \\ 
ii) using more complicated criteria for the data collection \\
iii)
finite horizon optimization where there has been much creative work in the robotics literature including probabilistic road maps and rapidly-exploring random trees \citep{karaman2011sampling} whose statistical properties are not yet clear.

We have further strengthened the discussion on communication for the AUV. In doing so, we relate the topics of autonomy and communication to the modeling and computational assumptions that applies to this paper. With the possibilities of wifi communication to a mother-ship at waypoints, more advanced modeling and computational routines are possible, of course with new challenges related to design and planning for when and where this communication should occur. This topic has caught much interest in control engineering \citep{zolich2019survey}, and statistical evaluations could contribute in this setting.

@@

\item{\revcom - Major 2:}\label{r3c3}

Figure 4: Why not show excursion probabilities for, say, myopic selection as well?

\reply: These plots are changed in the revised version.

In the accompanying code (\url{https://cedrictravelletti.github.io/Submarine/} )
one can visualize various sampling results and also display other outputs such as instance univariate excursion probabilities.

### 2020-06-27 main_new.tex
updates

% New Marcos.
\usepackage{amssymb}
\usepackage{hyperref}
\usepackage{cleveref}
\input{macros}

@@

% "New Title of the paper"
\title{Autonomous Oceanographic Data Collection. Learning Excursion Sets of Vector-valued Gaussian Random Fields} 
\runtitle{Autonomous Oceanographic Data Collection}

@@

\begin{keyword}
\kwd{Ocean Sampling}
\kwd{Excursion Sets}
\kwd{Gaussian Processes}
\kwd{Experimental Design}
\kwd{Autonomous robots}
\kwd{Adaptive Information Gathering}
\end{keyword}

@@

Other statistical work in the oceanographic domain include
\cite{wikle2013modern} focusing on hierarchical statistical models,
\cite{sahu2008space} studying spatio-temporal models for sea surface
temperature and salinity data and \cite{mellucci2018oceanic} looking
at the statistical prediction of features using an underwater glider.
In this paper the main focus is not on statistical modeling per se, but rather on statistical principles and computations underlying efficient data collection. We combine novel possibilities in marine
robotics with spatial statistics and experimental design to provide
useful AUV sampling designs.

@@

\begin{itemize}
\item Extending uncertainty reduction criteria to
  vector-valued cases.  
\item Closed-form expressions for the expected integrated Bernoulli
  variance (IBV) of the excursions in GRFs. 
\item Algorithms for myopic and multiple-step ahead sequential
  strategies for optimizing AUV sampling with respect to the mentioned
  criteria. 
\item Replicable experiments on synthetic cases with accompanying
  code~\footnote{\url{https://github.com/CedricTravelletti/MESLAS}}
\item Results from full-scale field trials running myopic strategies onboard an AUV for the characterization of a river plume. 
\end{itemize}

@@

We now turn to an example involving a full-fledged GRF. The statistical model we consider has a linear trend
\begin{align*}
\mu(s)=\mathbb{E}\left[\begin{pmatrix}
Z_{\spatloc, 1}\\ Z_{\spatloc, 2}
\end{pmatrix}\right] &= \beta_0 + \beta_1 \spatloc,
\end{align*}
with $\beta_0$ a two dimensional vector and $\beta_1$ a $2\times 2$ matrix. In our examples, we only consider separable covariance models;
\begin{align*}
\textrm{Cov}\left(Z_{\spatloc, i}, Z_{v, j}\right) &= k(\spatloc, v) \gamma(i, j),~ \gamma(i, j) = \begin{cases} \sigma_i^2,~ i=j\\
   \gamma \sigma_i \sigma_j,~i\neq j,
        \end{cases}
\end{align*}
where an isotropic Mat\'{e}rn 3/2 kernel $(1+\eta h)\exp (-\eta h)$ is used, for Euclidean distance $h$.
In the accompanying Python
examples taking place within the MESLAS toolbox~\footnote{\url{https://github.com/CedricTravelletti/MESLAS}}, these
modeling assumptions can however be relaxed to anisotropic covariance and
changing variance levels across the spatial domain. Both extensions are relevant for the setting with river
plumes, but in practice this requires more parameters to be specified. With extensive satellite data or prior knowledge from high-resolution ocean models, one could also possibly fit more complex multivariate
spatial covariance functions
\citep{gneiting2010matern,genton2015cross}, but that is outside the
scope of the current paper.

@@

\section{Sequential designs and heuristic path planning}\label{sec:heuristics}

@@

\subsection{Look-ahead Trajectory Planning}\label{sec:LA}

@@

\subsection{Simulation study}
\label{sec:simulations}

@@

This work builds on a multidisciplinary effort combining statistical
methods with robotic surveying techniques for oceanographic
applications. We show how observation practices can gain efficiency
and accuracy from the development of statistical techniques for
spatial monitoring. We further demonstrate the opportunities available
for real-time multivariate spatial data gathering and analysis on
autonomous platforms, which statisticians can exploit to create
general-purpose toolkits for similar applications.

In particular, we derive and show results for characterizing phenomena
connected to the properties of water masses. The characterization of
uncertainties in random sets is extended in the vector-valued case with new results for the
expected integrated Bernoulli variance reduction achieved by spatial
sampling designs. This is provided in semi-analytical form for the situation
with a static design, and then extended to the adaptive situation. The
sequential derivations provide new insights into efficient
applications of adaptive data collection, as demonstrated in our
application.

The case study considers the upper water column in the river plume,
represented by a two dimensional grid. Extensions to three-dimensional
domains are not methodologically different, but the operation must
likely approximate calculations by concentrating numerical integration efforts on terms in the
vicinity of the autonomous vehicle \citep{fossum18b}. We did not
consider any temporal effects, which would be relevant on a larger
time scale. We consider the extension to spatio-temporal modeling as
future work, and envision that advection-diffusion equations could be
useful in this kind of modeling
\citep{sigrist2015stochastic,richardson2017sparsity}. For more complex
oceanographic phenomena, the methods will need to be extended to
non-Gaussian phenomena, possibly feature-based mixtures of Gaussian
processes which could still be run onboard and augmented by dynamical
models. If high-resolution ocean model or satellite data are available, it is hopefully possible to fit such more complex statistical models.

The spatial-statistical design criterion building on random sets is
relevant in our setting with different water properties.  We show
mathematical generality beyond the expected integrated Bernoulli
variance, for instance that of volume uncertainties which is possibly
more relevant, but requires more computational resources.  Such
criteria could be particularly useful in other oceanographic settings
related to mapping of algal-blooms, anoxic zones or open water fronts
\cite{costa19}.  Of course, other criteria are also relevant for
instance, hybrid or multi-attribute criteria that could balance goals
of exploration and exploitation in this situation. Equally, such
techniques have significant use cases in downstream decision-making,
with policy makers and regulators who need to make difficult decisions
related to fish farming or other marine resource operations. Value of
information analysis \citep{Eidsvik:15} could be used to evaluate
whether information is likely to result in improved decision-making,
in such a context.  We also foresee opportunities related to design of
experiments for multivariate processes using our notion of generalized
locations. For instance, this is likely to be useful for
internet-of-things applications or computer experiments where some
observations or evaluations are rather inexpensive, while others 
must only be done when they are really valuable.

In our context the myopic strategy performs well, and for
computational reasons we did not go in depth on the dynamic program
solution. There has been much creative work on finite horizon
optimization in the robotics literature including probabilistic road
maps and rapidly-exploring random trees \citep{karaman2011sampling},
but their statistical properties are not yet clear.  
In some cases it is also limiting to use the waypoint graph, and it would be beneficial to allow more continuous updates and navigation at the highest frequency allowed by the AUV.
It is equally
interesting to explore the additional flexibility that can be gained
by having multiple vehicles co-temporally exploring a spatial or
spatio-temporal domain \citep{ferreira2019advancing}. Such an approach
would enable concurrent sampling in different parts of the space, or
opportunities to move in parallel to best capture the excursion set.
The sometimes conflicting topics of autonomy and communication have
caught much interest in control engineering lately
\citep{zolich2019survey}, and statistical evaluations could contribute
in the future. In our context, autonomy is highlighted, and this is
reflected in our modeling and computational assumptions. With the
possibilities of wifi communication to a mother-ship at waypoints,
more advanced modeling and computational routines are of course
possible, but with new challenges related to design and planning for
when and where the communication should occur.

@@

TOF acknowledges support from the Centre for Autonomous Marine
Operations and Systems (AMOS)\footnote{\url{https://www.ntnu.edu/amos}}, Center of
Excellence, project number 223254, and the Applied Underwater Robotics Labortatory (AURLab). CT and DG acknowledge
support from the Swiss National Science Foundation, project number
178858. JE and KR acknowledge support from Norwegian
research council (RCN), project number 305445. DG would like to aknowledge support of Idiap Research Institute, where he had is main affiliation when the first version of this work was submitted. The authors would also like to thank Niklas Linde (University of Lausanne) for pointing out the relevance of revisiting the proposed approach for estimating volumes of mixing zones, e.g. in hydrology.

>>>>

% New Marcos.
\usepackage{amssymb}
\usepackage{hyperref}
\usepackage{cleveref}
\usepackage{xcolor}
\input{macros}

@@

% "New Title of the paper"
\title{Autonomous Oceanographic Data Collection. Learning Excursion
  Sets of Vector-valued Gaussian Random Fields \\or\\
Learning Excursion Sets of Vector-valued Gaussian Random Fields for
Autonomous Ocean Sampling
}

\runtitle{Autonomous Oceanographic Data Collection}

@@

\begin{keyword}
\kwd{Excursion Sets}
\kwd{Gaussian Processes}
\kwd{Experimental Design}
\kwd{Autonomous robots}
\kwd{Ocean Sampling}
\kwd{Adaptive Information Gathering}
\end{keyword}

@@

Other statistical work in the oceanographic domain include
\cite{wikle2013modern} focusing on hierarchical statistical models,
\cite{sahu2008space} studying spatio-temporal models for sea surface
temperature and salinity data and \cite{mellucci2018oceanic} looking
at the statistical prediction of features using an underwater glider.
In this work the main focus is not on statistical modeling per se, but
rather on statistical principles and computation underlying efficient
data collection. We combine novel possibilities in marine robotics
with spatial statistics and experimental design to provide useful AUV
sampling designs.

@@

\begin{itemize}
\item Extending uncertainty reduction criteria to
  vector-valued cases.  
\item Closed-form expressions for the expected integrated Bernoulli
  variance (IBV) of the excursions in GRFs. 
\item Algorithms for myopic and multiple-step ahead sequential
  strategies for optimizing AUV sampling with respect to the mentioned
  criteria. 
\item Replicable experiments on synthetic cases with accompanying
  code
\item Results from full-scale field trials running myopic strategies onboard an AUV for the characterization of a river plume. 
\end{itemize}

@@

We now turn to an example involving a full-fledged GRF. The statistical model we consider has a linear trend
\begin{align*}
\mu(s)=\mathbb{E}\left[\begin{pmatrix}
Z_{\spatloc, 1}\\ Z_{\spatloc, 2}
\end{pmatrix}\right] &= \beta_0 + \beta_1 \spatloc,
\end{align*}
with $\beta_0$ a two dimensional vector and $\beta_1$ a $2\times 2$ matrix. In our examples, we only consider separable covariance models;
\begin{align*}
\textrm{Cov}\left(Z_{\spatloc, i}, Z_{v, j}\right) &= k(\spatloc, v) \gamma(i, j),~ \gamma(i, j) = \begin{cases} \sigma_i^2,~ i=j\\
   \gamma \sigma_i \sigma_j,~i\neq j,
        \end{cases}
\end{align*}
where an isotropic Mat\'{e}rn 3/2 kernel $(1+\eta h)\exp (-\eta h)$ is
used, for Euclidean distance $h$.  In the accompanying Python examples
taking place within the MESLAS
toolbox~\footnote{\url{https://github.com/CedricTravelletti/MESLAS}},
these modeling assumptions can however be relaxed to anisotropic
covariance and changing variance levels across the spatial
domain. Both extensions are relevant for the setting with river
plumes, but in practice this requires more parameters to be
specified. With extensive satellite data or prior knowledge from
high-resolution ocean models, one could also possibly fit more complex
multivariate spatial covariance functions
\citep{gneiting2010matern,genton2015cross}, but that is outside the
scope of the current work.

@@

\section{Sequential designs and heuristic path planning}
\label{sec:heuristics}

@@

\subsection{Look-ahead Trajectory Planning}
\label{sec:LA}

@@

\subsection{Simulation studies}
\label{sec:simulations}

@@

This work builds on a multidisciplinary effort combining statistical
methods with robotic sampling for oceanographic applications. We show
how observation practices can gain efficiency and accuracy from
statistical techniques for spatial monitoring and demonstrate the the
need for real-time multivariate spatial sampling on autonomous
platforms.% , which statisticians can exploit to create general-purpose
% toolkits for similar applications.

In particular, we derive and show results for a real-world domain
characterizing water mass properties. The characterization of
uncertainties in random sets is extended in the vector-valued case
with new results for the expected integrated Bernoulli variance
reduction achieved by spatial sampling designs. This is provided in
semi-analytical form for static designs, and then extended to the
adaptive situations. The sequential derivations provide new insights
into efficient applications of adaptive data collection, as
demonstrated in our application.

The case study considers the upper water column in the river plume,
represented by a two dimensional grid. Extensions to three-dimensional
domains are not methodologically different, but likely approximate
calculations by concentrating numerical integration on terms in the
vicinity of the autonomous vehicle \citep{fossum18b}. While we did not
consider any temporal effects, which would be relevant on a larger
time scale, we do consider the extension to spatio-temporal modeling
and envision that advection-diffusion equations could be useful in
such modeling
\citep{sigrist2015stochastic,richardson2017sparsity}. For more complex
oceanographic phenomena, the methods will need to be extended to
non-Gaussian phenomena, possibly feature-based mixtures of Gaussian
processes which could still be run onboard and augmented by dynamical
models. If high-resolution ocean model or satellite data are
available, it is hopefully possible to fit such more complex
statistical models. \kc{'Fit' as in modeling fit, or fit for
  augmentation with dynamical models? But also note ocean models are
  unlikely to be able to run onboard, so a simple phrase should be
  added to that effect.}

The spatio-statistical design criterion building on random sets is
relevant in our setting with different water properties.  We show
mathematical generality beyond the expected integrated Bernoulli
variance, for instance, that of volume uncertainties which is possibly
more relevant, but one that requires more computational resources.
Such criteria could be particularly useful in other oceanographic
settings related to mapping of algal-blooms, anoxic zones or open
water fronts \cite{costa19}.  Other criteria could also be relevant,
for instance, hybrid or multi-attribute criteria that could balance
goals of exploration and exploitation in this situation. Equally, such
techniques have significant use cases in downstream decision-making,
with policy makers and regulators who need to make difficult decisions
related to aquaculture or other marine resources. Value of information
analysis \citep{Eidsvik:15} could be used to evaluate whether
information is likely to result in improved decision-making, in such a
context.  We also foresee opportunities related to design of
experiments for multivariate processes using our notion of generalized
locations.

In our context the myopic strategy performs well, and due to
computational constraints we did not go in depth on the dynamic
programming solutions. There has been much work on finite horizon
optimization in the robotics literature including probabilistic road
maps and rapidly-exploring random trees \citep{karaman2011sampling},
but their statistical properties are unclear.  In some cases it is
also limiting to use a waypoint graph, and would be beneficial to
allow more continuous updates and navigation at the highest frequency
possible given limitations being onboard an AUV.  It is equally
interesting to explore the additional flexibility that can be gained
by having multiple vehicles co-temporally exploring a spatial or
spatio-temporal domain \citep{ferreira2019advancing}. Such an approach
would enable concurrent sampling in different parts of the space, or
opportunities to move in parallel to best capture the excursion set.
The value of information related to when and what to communicate (to
shore or to other vehicles) is also an interesting thrust for research
and likely to be useful for internet-of-things applications or
computer experiments where some observations or evaluations are rather
inexpensive, while others must only be done when they are really
valuable.

% The sometimes conflicting topics of autonomy and communication have
% caught much interest in control engineering lately
% \citep{zolich2019survey}, and statistical evaluations could contribute
% in the future. In our context, autonomy is highlighted, and this is
% reflected in our modeling and computational assumptions. With the
% possibilities of wifi communication to a research vessel at waypoints,
% more advanced modeling and computational routines are possible, but
% with new challenges related to design and planning for when and where
% the communication should occur.

@@

TOF acknowledges support from the Centre for Autonomous Marine
Operations and Systems (AMOS), Center of Excellence, project number
223254, and the Applied Underwater Robotics Labortatory (AURLab). CT
and DG acknowledge support from the Swiss National Science Foundation,
project number 178858. JE and KR acknowledge support from Norwegian
research council (RCN), project number 305445. DG would like to
aknowledge support of Idiap Research Institute, his primary
affiliation in an early version of this work.  The authors would also
like to thank Niklas Linde of the University of Lausanne for providing
constructive feedback on a version of the manuscript.

%  for pointing out the relevance of revisiting the proposed
% approach for estimating volumes of mixing zones, e.g. in hydrology.
